Nail: A practical interface generator for data formats

### 2014-11-24 paper/login/design.tex
login

The body of the \texttt{dnspacket} rule is a structure, which contains any number of fields enclosed
between  curly braces. Each field in the structure is parsed in sequence, and represented as a
structure to the programmer.  Contrary to other programming languages, Nail does not have a special
keyword for structs. We also reverse the usual structure-field syntax: \texttt{id uint1} is a field
called \texttt{id} with type \texttt{uint1}\footnote{This is a feature from Go, and makes nested
  types more readable and the parser for Nail itself easier to write.}. Often, Nail grammars have
structures with just one non-constant field, for example when parsing a fixed header. To prevent
this from cluttering the data type, we support an alternative form of the structure combinator,
using angle brackets, that has just one unnamed, non-constant field, which is represented directly
in the data type.

@@

\subsection{Input streams and transformations}
\label{s:transforms}

Traditional parsers handle input one symbol at a time, from beginning to end.
However, real-world formats often require non-linear parsing. Offset fields require a parser to move
to a different position in the input, possibly backwards. Size fields require the parser to stop
processing before the end of input has been reached, and perhaps resume executing a parent parser.
Other cases, such as compressed data, require more complicated processing on parts of the input
before it can be handled.

Nail introduces two concepts to handle these challenges, \emph{streams} and \emph{transformations}. 
Streams represent a sequence of bytes that contain some external format. The parsers and generators
that Nail generates always operate on an implicit stream named \cc{\$current} that they process front to
back, reading input or appending output.
Grammars can use the \cc{apply} combinator to parse a stream or generate output into it.

Streams are passed as arguments to a rule or defined within the grammar through \emph{transformations}.
The current stream is always passed as an implicit parameter.

@@

Typically, the top level of most grammars is a rule that takes only a single stream, which may then
be broken up by various transformations and passed to sub-rules, which eventually parse various linear
fragment streams. Upon parsing, these fragment streams are generated and then combined by the
transforms.

 To reduce both programmer effort and the risk of unsafe operations, Nails provides implementations of
transformations for many common features, such as checksums, size, and offset fields. Furthermore,
Nail provides library functions that can be used to safely operate on streams, such as splitting and
concatenation. Nail implements streams as iterators, so they can share underlying buffers and can be
efficiently duplicated and split.

@@

\paragraph{Compressed data.}
Encoded, compressed, or encrypted data can be handled transparently by writing a custom
transformation that transforms a coded stream into one that can be parsed by a Nail grammar and vice
versa. This transformation must be carefully written to not have bugs.

For example, DNS uses label compression to reduce the overhead of including each domain name
multiple times in a DNS reply. If a domain name suffix is repeated, it may be replaced by a two-bit
marker followed by a  14-bit offset into the packet, indicating the position where that suffix was
previously encoded.  We use a transform in line 35
of our DNS grammar to handle DNS label compression; the signatures of the two transform functions are shown
in Figure~\ref{fig:dns-xform}.  When parsing a packet, this transform
decompresses the DNS label stream.
When generating a packet, this transform receives the current suffix as
an input, and scans the packet so far for previous occurrences and inserts the offset.

@@

\paragraph{Offsets.}
A built-in transformation for handling offset fields, which is invoked as follows:
\cc{\$fragment transform offset\_u32(\$current, \cc{@}offset)}. 
This transformation corresponds to two functions for parsing and
generation, as shown in Figure~\ref{fig:xform-sig}. It defines a new stream \cc{\$fragment} that can
be used to parse data at the offset contained in \cc{@offset}, by
using \cc{apply \$fragment some_parser}.

>>>>

The body of the \texttt{dnspacket} rule is a structure, which contains any number of fields enclosed
between  curly braces. Each field in the structure is parsed in sequence, and represented as a
structure to the programmer.  Contrary to other programming languages, Nail does not have a special
keyword for structs. We also reverse the usual structure-field syntax: \texttt{id uint1} is a field
called \texttt{id} with type \texttt{uint1}. Often, Nail grammars have
structures with just one non-constant field, for example when parsing a fixed header. Nail supports
this with an alternative form of the structure combinator,
using angle brackets, that contains one unnamed, non-constant field, which is represented directly
in the datatype, without introducing another layer of indirection, as shown on line 40.

@@

\subsection{Input streams and transformations}
\label{s:transforms}
Formats have

Traditional parsers handle input one symbol at a time, from beginning to end.
However, real-world formats often require non-linear parsing. Offset fields require a parser to move
to a different position in the input, possibly backwards. Size fields require the parser to stop
processing before the end of input has been reached, and perhaps resume executing a parent parser.
Other cases, such as compressed data, require more complicated processing on parts of the input
before it can be handled.
Streams represent a sequence of bytes that contain some external format. The parsers and generators
that Nail generates always operate on an implicit stream named \cc{\$current} that they process front to
back, reading input or appending output.
Grammars can use the \cc{apply} combinator to parse a stream or generate output into it. 
Streams are passed as arguments to a rule or defined within the grammar through \emph{transformations}.
The current stream is always passed as an implicit parameter.

There are many ways conceptual streams could be encoded in a data format --
A tool like Nail needs to support all of them. This makes data description languages like
PADS~\cite{Fisher:2005:PDL:1064978.1065046} contain not just a kitchen sink, but a kitchen store
full of features -- and a language that can handle all possible format oddities will be a general
purpose language.  Nail instead of allows you to plug in your own code as \emph{transformations},
with a small, but powerful camping kitchen included that will handle most formats.

@@

Typically, grammars start out with a single stream, which may then
be broken up by various transformations and passed to sub-rules, which eventually parse various linear
fragment streams. Upon parsing, these fragment streams are generated and then combined by the
transforms.

 To make your life easier and your code safer, Nails provides implementations of
transformations for many common features, such as checksums, size, and offset fields. Furthermore,
Nail provides library functions that can be used to safely operate on streams, such as splitting and
concatenation. Nail implements streams as iterators, so they can share underlying buffers and can be
efficiently duplicated and split.

@@

\paragraph{Compressed data.}
For example, DNS uses label compression to reduce the overhead of including each domain name
multiple times in a DNS reply. If a domain name suffix is repeated, it may be replaced by a two-bit
marker followed by a  14-bit offset into the packet, indicating the position where that suffix was
previously encoded.  We use a transform in line 35
of our DNS grammar to handle DNS label compression; the signatures of the two transform functions are shown
in Figure~\ref{fig:dns-xform}.  When parsing a packet, this transform
decompresses the DNS label stream.
When generating a packet, this transform receives the current suffix as
an input, and scans the packet so far for previous occurrences and inserts the offset.

@@

\paragraph{Offsets.}
We provide a  transformation for handling offset fields, which is invoked as follows:
\cc{\$fragment transform offset\_u32(\$current, \cc{@}offset)}. 
This transformation corresponds to two functions for parsing and
generation, as shown in Figure~\ref{fig:xform-sig}. It defines a new stream \cc{\$fragment} that can
be used to parse data at the offset contained in \cc{@offset}, by
using \cc{apply \$fragment some_parser}.

### 2014-11-24 paper/login/eval.tex
Login
login

\paragraph{Performance}

To evaluate whether Nail-based parsers are compatible with good
performance, we compare the performance of our DNS server to that of ISC
BIND 9 release 9.9.5~\cite{bind9}, a mature and widely used DNS server.
We simulate a load resembling that of an authoritative name server, generating a random zone file
and a random sequence of queries, with 10\% non-existent domains. We repeated this sequence of
queries for one minute against both DNS servers. The results are shown in Figure~\ref{fig:perf-dns}, and demonstrate that
our Nail-based DNS server can achieve higher performance and lower latency
than BIND\@.  Although BIND is a more sophisticated DNS server,
and implements many features that are not present in our Nail-based DNS
server and that allow it to be used in more complicated configurations, 
we believe our results demonstrate that Nail's parsers are not
a barrier to achieving good performance.
\paragraph{ZIP.}

@@

In our full paper~\cite{bangert:nail-osdi14}, we present a study of 15 ZIP parsing bugs.
11 of these vulnerabilities involved memory corruption during input handling, which generated code is immune to
memory corruption attacks by design.
More interestingly, Nail also protects against parsing inconsistency vulnerabilities like the four others we studied.
Nail grammars explicitly encode duplicated information such as the redundant length fields in ZIP
that caused a vulnerability in the Python ZIP library. The other three vulnerabilities exist
because multiple implementations of the same protocol disagree on some inputs. Hand-written protocol
parsers are not very reusable, as they build application-specific data structures and are tightly
coupled to the rest of the code. Nail grammars, however, can be re-used between applications,
avoiding protocol misunderstandings.

>>>>

To evaluate whether Nail-based parsers are compatible with good
performance, we compare the performance of our DNS server to that of ISC
BIND 9 release 9.9.5~\cite{bind9}, a mature and widely used DNS server.
We simulate a load resembling that of an authoritative name server, generating a random zone file
and a random sequence of queries, with 10\% non-existent domains. We repeated this sequence of
queries for one minute against both DNS servers. We found that our DNS server is approximately three
times faster than BIND. Although BIND is a more sophisticated DNS server,
and implements many features that are not present in our Nail-based DNS
server and that allow it to be used in more complicated configurations, 
we believe our results demonstrate that Nail's parsers are not
a barrier to achieving good performance.
\paragraph{ZIP.}

@@

In our full paper~\cite{bangert:nail-osdi14}, we present a study of 15 ZIP parsing bugs.
11 of these vulnerabilities involved memory corruption during input handling, which generated code
is immune to by design.
More interestingly, Nail also protects against parsing inconsistency vulnerabilities like the four others we studied.
Nail grammars explicitly encode duplicated information such as the redundant length fields in ZIP
that caused a vulnerability in the Python ZIP library. The other three vulnerabilities exist
because multiple implementations of the same protocol disagree on some inputs. Hand-written protocol
parsers are not very reusable, as they build application-specific data structures and are tightly
coupled to the rest of the code. Nail grammars, however, can be re-used between applications,
avoiding protocol misunderstandings.

### 2014-11-24 paper/login/intro.tex
login

\begin{quote}
  I am ZIP, file of files.
  Parse me, ye mighty and despair.

@@

However, existing tools are not powerful enough to handle some complixities formats such as PDF
and ZIP, so programmers still have to hand-write code to handle the most complicated and therefore
most error-prone formats. For example, existing parser generators cannot conveniently represent size or offset fields and more complex features, such as data compression or checksums cannot be expressed at all.

Furthermore, some parser generators are cumbersome to use when parsing binary data for several
reasons. First, many parser generators don't directly data structures, but call semantic actions
that the programmer writes to construct a data structure. Therefore, programmers must
describe the format up to three times -- in the grammar, the data structure and the semantic
actions. Second, most parser generators only address parsing inputs, so programmers have to manually
construct outputs. Some parser generators, such as 
Boost.Spirit~\cite{boost-spirit} allow generating output, but require the programmer to write another set of semantic actions.

We address these challenges with Nail, a new parser generator for binary formats.
First, Nail grammars describe not only a format, but also a data type to represent it within the program.
This removes the semantic actions and type declarations that programmers have to write with existing
parser generators. It also forces the programmer to clearly separate syntactic validation and semantic processing.
Second, Nail does not just
parse input into an internal representation, but also generates output from the internal representation,
without requiring the programmer to write additional code.

Third, Nail introduces two abstractions, \emph{dependent fields} and
\emph{transformations}, to elegantly handle problematic structures,
such as offset fields or checksums.  Dependent fields capture fields
in a protocol whose value depends in some way on the value or layout
of other parts of the format; for example, offset or length fields,
which specify the position or length of another data structure, fall
in this category.  Transformations allow the programmer to escape the
generated code to modify the raw data and interact with dependent fields
in a controlled manner.

>>>>

\begin{quote}
  I am ZIP, file of files.
  Parse me, ye mighty and drop a shell.

@@

However, if you wanted to use existing tools to parse PDF
or ZIP, you would soon find that they cannot handle the complicated -- and therefore most
error-prone-- aspects of such formats-- so you'd still have to
hand-write the riskiest bits of code. For example, existing parser generators cannot conveniently represent size or offset fields and more complex features, such as data compression or checksums cannot be expressed at all.

Furthermore, some parser generators are cumbersome to use when parsing binary data for several
reasons. First, many parser generators don't directly data structures, but call semantic actions
that the you write to construct a data structure. Therefore,you must
describe the format up to three times -- in the grammar, the data structure and the semantic
actions. Second, most parser generators only address parsing inputs, so you'd have to manually
construct outputs. Some parser generators, such as 
Boost.Spirit~\cite{boost-spirit} allow generating output, but require you to write another set of semantic actions.

We address these challenges with Nail, a new parser generator for binary formats.
First, Nail grammars describe not only a format, but also a data type to represent it within the program.
Therefore you don't have to write semantic actions and type declarations and you can no longer syntactic validation and semantic processing.
Second, Nail will also take care of generating output from the same data type, without requiring you
to write more risky code or think about the format again.

Third,  Nail introduces two abstractions, \emph{dependent fields} and
\emph{transformations}, to elegantly handle problematic structures,
such as offset fields or checksums.  Dependent fields capture fields
in a protocol whose value depends in some way on the value or layout
of other parts of the format; for example, offset or length fields,
which specify the position or length of another data structure, fall
in this category.  Transformations allow you to escape the
generated code to modify the raw data and interact with dependent fields
in a controlled manner.
